Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision

### 2023-04-19 Milestone_presentation_2/Milestone_presentation_2/Milestone_presentation_2.Rmd
fix bib

.float-left[

They are **very good at detecting violations of model assumptions**, but there are some **unexpected usages**.

.float-left[

- Most residual-based tests for particular types of departure are also sensitive to **other types of departures**, resulting in a phenomenon known as the "**Type III error**" (Cook and Weisberg 1982).

.float-left[

- Other problems such as **outliers** can trigger the rejection (Cook and Weisberg 1999).

```{r}
set.seed(10086)
# Base model
mod <- poly_model(include_z = FALSE, sigma = 0.25)
# Base data
dat <- mod$gen(300)
bind_rows(
  dat %>%
    mutate(type = "(A) Good residuals"),
  mod$set_prm("include_z", TRUE)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(B) Non-linearity"),
  heter_model(b = 64)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(C) Heteroskedasticity"),
  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
    gen(300, computed = select(dat, x)) %>%
    mutate(type = "(D) Non-normality")
) %>%
  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
  facet_wrap(~type, ncol = 2, scales = "free") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) +
  xlab("Fitted values") +
  ylab("Residuals")
```

All the hypothesis tests **correctly reject** the correspoding model defects, and **correctly not reject** the structure in residual plot (A).

BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for**.

```{r}
set.seed(10086)
# Base model
mod <- poly_model(include_z = FALSE, sigma = 0.25)
# Base data
dat0 <- mod$gen(300)
# Replicate data in Figure 1
dat1 <- mod$set_prm("include_z", TRUE)$gen(300, computed = select(dat0, x, e))
dat2 <- heter_model(b = 64)$gen(300, computed = select(dat0, x, e))
dat3 <- heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
               gen(300, computed = select(dat0, x))
dat_list <- list(dat0, dat1, dat2, dat3)
table_dat <- data.frame(plot = c("A", "B", "C", "D"), 
                        model = c("None", 
                                  "Non-linearity", 
                                  "Heteroskedasticity", 
                                  "Non-normality"),
                        r = map_dbl(dat_list, 
                                    ~POLY_MODEL$test(.x, 
                                                     test = "RESET", 
                                                     power = 2:4)$p_value),
                        b = map_dbl(dat_list, 
                                    ~HETER_MODEL$test(.x)$p_value),

                        # SW test has not been included by visage
                        s = map_dbl(dat_list, 
                                    ~shapiro.test(.x$.resid)$p.value)) %>%

  # 3 decimal points
  mutate(across(r:s, ~ format(round(.x, digits = 3), nsmall = 3))) %>%

  # Italic if reject
  mutate(across(r:s, ~ kableExtra::cell_spec(.x, italic = as.numeric(.x) <= 0.05, color = ifelse(as.numeric(.x) <= 0.05, "red", "black"))))
  table_dat %>%
  knitr::kable(col.names = c("Plot", 
                             "Departures", 
                             "RESET", 
                             "BP", 
                             "SW"),
               align = "llrrr",
               escape = FALSE,
               format = "html",
               booktabs = TRUE)
```

@@

This framework is called **visual inference**. It was introduced by Buja, et al. (2009) as an inferential framework to extend confirmatory statistics to visual discoveries.

A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) containing data consistent with the null hypothesis.

@@

### (uniform fitted values only)

@@

.center[
# ⚖️Results - Power Comparison of the Tests⚖️
]

@@

.center[
# ⚖️Results - Comparison of Test Decisions⚖️
]

@@

.center[
# ⚖️Results - Effect of Amount of Non-linearity⚖️
]

@@

.center[
# ⚖️Results - Effect of Shape of Heteroskedasticity⚖️
]

@@

>>>>

@@

This framework is called **visual inference** (Buja, et al. 2009).

A **lineup** consists of $m$ randomly placed plots, where one plot is the **data plot** and the remaining $m − 1$ plots (**null plots**) containing data **consistent with the null hypothesis**.

@@

@@

.center[
# ⚖️Results - Power Comparison of the Tests⚖️

@@

.center[
# ⚖️Results - Comparison of Test Decisions⚖️

@@

.center[
# ⚖️Results - Effect of Amount of Non-linearity⚖️

@@

.center[
# ⚖️Results - Effect of Shape of Heteroskedasticity⚖️

@@

.font-70-p[

Belsley, David A, Edwin Kuh, and Roy E Welsch. 1980. Regression diagnostics: Identifying influential data and sources of collinearity. John Wiley & Sons.

Breusch, T. S., and A. R. Pagan. 1979. “A Simple Test for Heteroscedasticity and Random Coefficient Variation.” Econometrica 47 (5): 1287–1294.

Buja, Andreas, Dianne Cook, Heike Hofmann, Michael Lawrence, Eun-Kyung Lee, Deborah F. Swayne, and Hadley Wickham. 2009. “Statistical inference for exploratory data analysis and model diagnostics.” Philosophical Transactions of the Royal Society A: Mathematical, Physical and Engineering Sciences 367 (1906): 4361–4383.

Cook, R Dennis, and Sanford Weisberg. 1982. Residuals and influence in regression. New York: Chapman and Hall.

Cook, R Dennis, and Sanford Weisberg. 1999. Applied regression including computing and graphics. John Wiley & Sons.

Draper, Norman R, and Harry Smith. 1998. Applied regression analysis. Vol. 326. John Wiley & Sons.

Jarque, Carlos M, and Anil K Bera. 1980. “Efficient tests for normality, homoscedasticity and serial independence of regression residuals.” Economics letters 6 (3): 255–259.

Kullback, Solomon, and Richard A Leibler. 1951. “On information and sufficiency.” The annals
of mathematical statistics 22 (1): 79–86.

Ramsey, J. B. 1969. “Tests for Specification Errors in Classical Linear Least-Squares Regression
Analysis.” Journal of the Royal Statistical Society. Series B (Methodological) 31 (2): 350–
371.

Shapiro, Samuel Sanford, and Martin B Wilk. 1965. “An analysis of variance test for normality (complete samples).” Biometrika 52 (3/4): 591–611.

VanderPlas, Susan, Christian R ̈ottger, Dianne Cook, and Heike Hofmann. 2021. “Statistical significance calculations for scenarios in visual inference.” Stat 10 (1): e337.

White, Halbert. 1980. “A Heteroskedasticity-Consistent Covariance Matrix Estimator and a Direct Test for Heteroskedasticity.” Econometrica 48 (4): 817–838.
